---
title: "Final Project"
output: 
  prettydoc::html_pretty:
    theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### [HOME](https://chandlerhaut.github.io/) | [ABOUT ME](https://chandlerhaut.github.io/about_me/) | [CV](https://chandlerhaut.github.io/about_me/CV.pdf) | [FINAL PROJECT](https://chandlerhaut.github.io/Final_Project/)

## Background

The data used for this project is bat acoustic signals recorded on Wildlife Refuges throughout the Pacific Northwest. Acoustic data is recorded with a portable recording device and microphone, mounted upon a painters pole. These are place in riparian areas where bats are likely to obtain water and insects.  



```{r cars, echo=FALSE,message=FALSE}

library(tidyverse)
library(gapminder)
library(stringr)
library(leaflet)
library(sf)
library(shiny)
library(leaflet.extras)
library(htmltools)
library(ggmap)
library(knitr)
library(rmarkdown)

df <- read_csv('./R1_NABat_VettedObservations_NWRS2022.csv')

view(df)

bats <- df %>% 
  filter(!is.na(CommonName)) %>% 
  mutate(CommonName = str_replace(CommonName, "Townsend\x92s big-eared bat", 
                                  "Townsend big-eared bat")) %>% 
  select(-"organization_name",-"species_list",-"frame", -"project_name")

nb <- #Placed the below code in the origianl bats df to keep my SHIT together
  bats %>% 
  filter(!is.na(CommonName)) %>% 
  mutate(CommonName = str_replace(CommonName, "Townsend\x92s big-eared bat", 
                                  "Townsend big-eared bat"))


wdf <- 
nb %>% 
  filter(!is.na(confirmed)) %>% 
  group_by(CommonName, latitude, longitude, RefugeName) %>% 
  summarise(sum = sum(reviewed))

```

```{r}
print(df)
print(wdf)
```



## Interactive Map
```{r, echo=TRUE, message=FALSE}
wdf %>% 
  leaflet() %>% 
  addProviderTiles("Esri.WorldImagery") %>% 
  addCircleMarkers(radius = 10, label = ~paste0("Refuge =",htmlEscape(RefugeName)),
                   color = "blue", group = wdf$CommonName) %>% 
  addLayersControl(baseGroups = wdf$CommonName)
```

## How to get a map: 

```{r map, echo=TRUE, message=FALSE, results='hide'}
library(ggmap)

p <- get_map(location = c(lat = 43.618881, lon = -116.215019), 
                           zoom = 5, maptype = "terrain")

bbb <- 
  wdf %>% 
  filter(CommonName == "Big brown bat")

ggmap(p) +
  geom_density_2d(data = bbb, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Big Brown Bat")

```


## Distribution of: 

```{r,echo=FALSE, fig.show='hold', out.width="50%", message=FALSE}
bbb <- 
  wdf %>% 
  filter(CommonName == "Big brown bat")

ggmap(p) +
  geom_density_2d(data = bbb, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Big Brown Bat")


cm <- 
  wdf %>% 
  filter(CommonName == "California myotis")

ggmap(p) +
  geom_density_2d(data = cm, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("California myotis")

cb <- 
  wdf %>% 
  filter(CommonName == "Canyon bat") 

ggmap(p) +
  geom_density_2d(data = cb, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Canyon bat")

fm <- 
  wdf %>% 
  filter(CommonName == "Fringed myotis")

ggmap(p) +
  geom_density_2d(data = fm, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Fringed myotis")

hb <- 
  wdf %>% 
  filter(CommonName == "Hoary bat")

ggmap(p) +
  geom_density_2d(data = hb, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Hoary Bat")

lbm <- 
  wdf %>% 
  filter(CommonName == "Little brown myotis")

ggmap(p) +
  geom_density_2d(data = lbm, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Little brown myotis")

lem <- 
  wdf %>% 
  filter(CommonName == "Long-eared myotis") 

ggmap(p) +
  geom_density_2d(data = lem, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Long-eared myotis")

llm <- 
  wdf %>% 
  filter(CommonName == "Long-legged myotis") 

ggmap(p) +
  geom_density_2d(data = llm, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Long-legged Myotis")


shb <- 
  wdf %>% 
  filter(CommonName == "Silver-haired bat") 

ggmap(p) +
  geom_density_2d(data = shb, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Silver-haired bat")


tbeb<- 
  wdf %>% 
  filter(CommonName == "Townsend big-eared bat") 

ggmap(p) +
  geom_density_2d(data = tbeb, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Townsend Big-eared Bat")


wsfm<- 
  wdf %>% 
  filter(CommonName == "Western small-footed myotis") 

ggmap(p) +
  geom_density_2d(data = wsfm, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Western Small-footed Myotis")

ym <- 
  wdf %>% 
  filter(CommonName == "Yuma myotis") 

ggmap(p) +
  geom_density_2d(data = ym, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Yuma Myotis")
```





## Species of Greatest Concern:
```{r,echo=FALSE, fig.show='hold', out.width="50%", message=FALSE}
bftb <- 
  wdf %>% 
  filter(CommonName == "Brazilian free-tailed bat")

ggmap(p) +
  geom_density_2d(data = bftb, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Brazilian Free-tailed Bat")

pb <- 
  wdf %>% 
  filter(CommonName == "Pallid bat") 

ggmap(p) +
  geom_density_2d(data = pb, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Pallid Bat")

sb <- 
  wdf %>% 
  filter(CommonName == "Spotted bat") 

ggmap(p) +
  geom_density_2d(data = sb, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Spotted Bat")

wrb <- 
  wdf %>% 
  filter(CommonName == "Western red bat") 

ggmap(p) +
  geom_density_2d(data = wrb, aes(y=latitude, x=longitude), h = 2)+
  ggtitle("Western Red Bat")

```


## Conclusion:

Throughout the Pacific Northwest there are 16 species of bats, four (Brazilian Free-tailed Bat, Pallid Bat, Spotted Bat, Western Red Bat) which have a reduced range. One (Little Brown Myotis) species that resides in the Pacific Northwest is endangered but has a substantial presence. 

## Discussion: 

Comparing data with the Southwestern United States would be Acoustic recordings are a great way to identify species within an area but are not ideal to determine the population sizes for those species. Species with reduced ranges or protective status should be trapped through mist netting procedures



